---
title: "cBioAutomatedTools"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

cBioAutomatedTools package facilitates working with high-throughput data stored on <http://www.cbioportal.org/>. The official __Bioconductor package__ to obtain data in __R__ is `cgdsr`. To obtain data with this package, users have to pass a multi-step procidure. Besides, the index of cancers and thier subgroups might change frequently, which in turn, requires changing the R code. `cBioAutomatedTools` makes this procedure automated for __RNA-seq__, __microRNA-seq__, __microarray__ and __methylation__ data types. In addition, comparing the genetic data from different cancer studies/cancer study subgroups becomes much faster and easier.

The package consists of three functions: `data.availability()`, `process.data.for.cancers()` and `process.data.for.a.cancer()`.

## `data.availability()`

This function ,requires no input, scans all the cancer studies to find which of the __RNA-seq__, __microRNA-seq__, __microarray__ and __methylation__ data types exist for each cancer study. The result is stored in __global enviroment__ (user's workspace) as `available.data.types`. For convenience, the result are also stored as an excel file in the working directory entitled __Available Data Types.xlsx__.

    data.availability()

The following table shows how the output file looks like. Since the complete output is too large, only 5 rows are shown. 

|                                                                     | RNA-seq      |	MicroRNA-seq | Microarray (mRNA) | Microarray (miRNA) |	 Methylation
|:--------------------------------------------------------------------|:------------:|:-------------:|:------------------:|:-----------------:|:---------:|
| Acinar Cell Carcinoma of the Pancreas (Johns Hopkins, J Pathol 2014)| 	   -       |	     -       |       	-           |         -         |  	  -     |
| Acute Myeloid Leukemia (TCGA, NEJM 2013)                            | 	Available  |	     -       |       	-           |         -         |	Available |
| Acute Myeloid Leukemia (TCGA, Provisional)                          | 	Available  |       -       |       	-           |         -         |	Available	|
| Adenoid Cystic Carcinoma (FMI, Am J Surg Pathl. 2014)               | 	   -       |	     -       |       	-           |         -         |  	  -     |
| Adenoid Cystic Carcinoma (MDA, Clin Cancer Res 2015)                | 	   -       |	     -       |       	-           |         -         |  	  -     |




Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
